On Optimality of Finite Element Integration

### 2015-11-26 TOMS_Paper/assembly.tex
Start attacking, again, sharing elimination...

In this section, we study \textit{sharing elimination}, a transformation that aims to minimize the operation count by applying factorization, code motion and (if necessary) common sub-expression elimination to remove sharing. In generic problems, transformations of this sort usually require solving large combinatorial problems; for instance, to evaluate the impact of all possible factorization strategies. However, we show that since we restrict ourselves to a specific class of loops and expressions (described in Section~\ref{sec:lnopt}), we can formulate a simple algorithm ensuring quasi-optimality in general and optimality when expressions have a particular structure (typical of many finite element problems).

\subsubsection{Multilinear loop nest}
\label{sec:se-opln}
In the special case in which test and trial functions are in the same function space, there exists $w_1$ and $w_2$ such that $\alpha_{eij}^{w1} = \beta_{eik}^{w2}$. That is, common sub-expressions arise over different iteration spaces ($L_j$ and $L_k$). Both $\alpha$ and $\beta$ can be hoisted at the level of the closest common loop ($L_i$) and the common sub-expressions be replaced with suitable temporaries. This has been detailed in~\citeN{Luporini}.

@@

\subsubsection{Algorithm}
Algorithm~\ref{algo:opt} details the steps for performing sharing elimination on an abstract syntax tree representation of the loop nest. For simplicity, the algorithm neglects the pathological scenario described in Section~\ref{sec:se-opln}, in which the resulting operation count could only be very close to optimality when both test and trial functions are in the same function space. A constructive proof of optimality can be derived taking into account both the algorithm itself and the observations in Sections~\ref{sec:se-opln} and~\ref{sec:se-rln} (in which we have explored the entire factorization space).

>>>>

In this section, we study \textit{sharing elimination}, a transformation that aims to reduce the operation count by applying factorization, code motion and (if necessary) common sub-expression elimination to remove sharing. If optimality is the target and expressions are generally unstructured, transformations of this sort require solving large combinatorial problems, for instance to evaluate the impact of all possible factorization strategies. In our domain, however, expressions may be characterized by some properties. These properties hold either in general -- as inherent in finite element integration -- or in forms making ``special use'' of most common differential and tensor operators. We exploit them to build an algorithm that results in optimal or quasi-optimal sharing elimination depending on the problem to which it is applied. In Sections~\ref{sec:se-mln} and~\ref{sec:se-rln} we discuss these properties, while in Section~\ref{sec:sec:se-algo} we present the algorithm.

\subsubsection{Multilinear loop nest}
\label{sec:se-mln}
In the special case in which test and trial functions are in the same function space, there exists $w_1$ and $w_2$ such that $\alpha_{eij}^{w1} = \beta_{eik}^{w2}$. That is, common sub-expressions arise over different iteration spaces ($L_j$ and $L_k$). Both $\alpha$ and $\beta$ can be hoisted at the level of the closest common loop ($L_i$) and the common sub-expressions be replaced with suitable temporaries. This has been detailed in~\citeN{Luporini}.

@@

\subsubsection{Algorithm}
\label{sec:sec:se-algo}
Algorithm~\ref{algo:opt} details the steps for performing sharing elimination on an abstract syntax tree representation of the loop nest. For simplicity, the algorithm neglects the pathological scenario described in Section~\ref{sec:se-mln}, in which the resulting operation count could only be very close to optimality when both test and trial functions are in the same function space. A constructive proof of optimality can be derived taking into account both the algorithm itself and the observations in Sections~\ref{sec:se-mln} and~\ref{sec:se-rln} (in which we have explored the entire factorization space).
